\documentclass{article}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{latexsym}
\usepackage{times}
\usepackage{psfrag,epsfig,epsf}
\usepackage{graphics}
\usepackage{multirow}
\usepackage{fullpage}
\usepackage{verbatim}
\usepackage{fancyheadings}
\usepackage[T1]{fontenc}
\usepackage{arev}
\usepackage{subfigure}
\usepackage{url}
\usepackage[noline,noend,ruled,linesnumbered]{algorithm2e}
\usepackage{algpseudocode}
\linespread{1.02} 

\pagestyle{empty}

\addtolength{\topmargin}{-20pt}
\addtolength{\oddsidemargin}{-5pt}
\addtolength{\textwidth}{20pt}
\addtolength{\textheight}{50pt}

\newenvironment{myitem}{\begin{list}{$\bullet$}
{\setlength{\itemsep}{-0pt}
\setlength{\topsep}{0pt}
\setlength{\labelwidth}{0pt}
%\setlength{\labelsep}{0pt}
\setlength{\leftmargin}{10pt}
\setlength{\parsep}{-0pt}
\setlength{\itemsep}{0pt}
\setlength{\partopsep}{0pt}}}%
{\end{list}}

\begin{document}

\sloppy

\noindent \underline{CS 344: DESIGN AND ANALYSIS OF COMPUTER
  ALGORITHMS \hspace{1.6in} SPRING 2014}

\vspace{0.1in}

\begin{center}
{\bf {\large Homework 4}}\\
Searching for Paths on Graphs\\
\end{center}

\vspace{0.1in}

\noindent Deadline: April 20, 11:59pm.\\ 
Available points: 110. Perfect score: 100.\\

\begin{center}
{\bf Homework Instructions:}
\end{center}

\noindent {\bf Teams:} Homeworks should be completed by teams of
students - three at most. No additional credit will be given for
students that complete a homework individually. Please inform
Athanasios Krontiris {\bf if your team has changed from the previous
  assignments.} (email: ak979 AT cs.rutgers.edu).\\

\noindent {\bf Submission Rules:} Submit your solutions electronically
as a PDF document through \url{sakai.rutgers.edu}. Do not submit Word
documents, raw text, or hardcopies etc. Make sure to generate and
submit a PDF instead. Each team of students should submit only a
single copy of their solutions and indicate all team members on their
submission.  Failure to follow these rules will result in lower grade
in the assignment.\\

\noindent {\bf Late Submissions:} No late submission is allowed. If
you don't submit a homework on time, you get 0 points for that
homework.\\

\noindent {\bf Extra Credit for \LaTeX:} You will receive 10\% extra
credit points if you submit your answers as a typeset PDF (using
\LaTeX, in which case you should also submit electronically your
source code). Resources on how to use \LaTeX\ are available on the
course's website. There will be a 5\% bonus for electronically
prepared answers (e.g., on MS Word, etc.) that are not typeset.\\

\noindent {\bf 25\% Rule:} For any homework problem (same will hold
for exam questions), you can either attempt to answer the question, in
which case you will receive between 0 and 100\% credit for that
question (i.e., you can get partial credit), or you can write "I don't
know", in which case you receive 25\% credit for that question.
Leaving the question blank is the same as writing "I don't know." You
may get less than 25\% credit for a question that you answer
erroneously.\\

\noindent {\bf Handwritten Reports:} If you want to submit a
handwritten report, scan it and submit a PDF via Sakai. We will not
accept hardcopies. If you choose to submit handwritten answers and we
are not able to read them, you will not be awarded any points for the
part of the solution that is unreadable.\\

\noindent {\bf Precision:} Try to be precise. Have in mind that you
are trying to convince a very skeptical reader (and computer
scientists are the worst kind...) that your answers are correct.\\

\noindent {\bf Collusion, Plagiarism, etc.:} Each team of students
must prepare its solutions independently from other teams, i.e.,
without using common notes or worksheets with other students or trying
to solve problems in collaboration with other teams.  You must
indicate any external sources you have used in the preparation of your
solution. Do not plagiarize online sources and in general make sure
you do not violate any of the academic standards of the course, the
department or of the university (the standards are available through
the course's website). Failure to follow these rules may result in
failure in the course.\\

\newpage

\vspace{0.1in}

{\bf }

\noindent {\bf Problem 1: [25 points]} After the departure of Storm
Sandy, the small town of New Town suffered an electricity power
shortage. After this experience, the local electricity agent decided
to create plans to better handle future events. The idea is to send
out immediately after the power failure a truck to investigate the
damage along every street in the town. In order to recover the
electricity network as soon as possible, they want to find an
efficient tour so that the truck visits each street exactly once and
then goes back to the town headquarters. Suppose each street in New
Town is a two-way street. The task is to find whether there exists a
tour that visits each street only once and then goes back to the
starting point.\\

\noindent {\bf A.} Show that the problem has a solution if and only if
there are no dead ends and no $n$-way junctions in New Town, where $n$
is odd, i.e., each intersection in the town has an even degree.\\

\noindent {\bf B.} If such tour exists, please describe an efficient
algorithm to find the tour in linear time, i.e., in at most $O(|V| +
|E|)$ time, where $V$ is the set of junctions and $E$ is the set of
streets.\\

\noindent {\bf Problem 2: [25 points]} In a chemistry class, the
professor asks you to synthesize products using existing
chemicals. For instance, 2 parts of chemical A produces 1 part of
chemical B, or 3 parts of chemical C can give us 1 part of chemical
D. Lets assume that we have $n$ different chemicals that we can
mix. The production ratio given two chemicals is fixed and does not
change. These ratios are given in an $n \times n$ matrix, between all
the chemicals. In particular, using $x$ parts of chemical $i$ you will
receive $R_{ij} \cdot x$ parts of chemical $j$, or using $y$ parts of
chemical $j$ you will receive $R_{ji} \cdot y$ parts of chemical
$i$. Note that $R_{ij}$ is not necessarily an integer, it can be a
real number.\\

\noindent Because of the principle of mass conservation the ratios
have to satisfy the following requirement: $R_{ij} \cdot R_{ji} <
1$. This means that converting an amount of chemical $i$ to chemical
$j$ and then back to chemical $i$ should result in a smaller quantity
of chemical $i$ relatively to the original one.\\

\noindent {\bf A.} The Chemistry professor asks from you the
following. Given the $n \times n$ matrix, find the best sequence of
chemical reactions to produce the most of chemical $D$ given an
available quantity of chemical $A$. As mentioned before there are $n$
different chemicals you can use and the matrix specifies the
production ratios of all corresponding reactions. Provide an algorithm
for finding the best sequence. Argue about the running time of your
algorithm.\\

\noindent {\bf B.} One of the teams found out that using chemical
$i_1$ to produce chemical $i_2$, then chemical $i_3$, \dots, then
chemical $i_k$ and back to $i_1$ results in a bigger quantity of
chemical $i_1$ than the original one, i.e., $R_{i_1,i_2} \cdot
R_{i_2,i_3} \cdot \ldots \cdot R_{i_k,i_1} \geq 1$. Is this a problem
in the ratio matrix or the students did something wrong in their
calculations? Provide an efficient algorithm that can detect if such
an impossible sequence of chemical reactions seems feasible given the
$n \times n$ production ratio matrix. Argue about the running time of
your algorithm.\\

\noindent {\bf Problem 3: [25 points]} Consider a courier company with
a big network of transfer stations and trucks. You are responsible for
coordinating the routing of packages through the network. You have
access to and the database indicating the departure/arrival locations
and times for $m$ trucks among $n$ transfer locations.\\

\noindent You have to design an algorithm that receives a package's
origin and destination location (the location always corresponds to
one of the transfer stations), then output the list of trucks that can
allow the package to arrive at the destination as early as
possible. Note that in every intermediate transfer station, it takes
about an hour for a package to move from a truck to another. You need
to take this time into account.\\

\noindent Give an algorithm as efficient as possible to solve this
problem, argue its running time, and prove its correctness.\\

\noindent {\bf Problem 4: [35 points]} The clash of kings has come to an end in
the land of Westeros and the winds of winter have passed... The Grand
Maester has proposed a series of jousting tournaments to celebrate the
events. These tournaments will be used to decide who are the finest
knights in the Seven Kingdoms that should be selected as new members
of the Kingsguard.\\

\noindent Every town in Westeros wants to host such a tournament. Even
Pike at the Iron Islands wants to host a tournament, although the
ironborn are not known to hold any love for the peoples of the
mainland...\\

\noindent This posed a challenge for the Grand Maester. How should he
rank the knights given a large number of tournaments and different
outcomes at each tournament? He asked the advice of the Citadel and
selected the following strategy.\\

\noindent Each local Maester will be responsible to report a ranking
of the knights at each individual tournament. Then, it is possible to
define a preference score between two knights:
$$d(A,B) = \textrm{ \{ the number of tournaments knight A was ranked
  higher than knight B\} }$$ Obviously, if $d(A,B) > d(B,A)$, then
knight $A$ is preferred as a member of the Kingsguard over candidate
$B$.\\

\noindent Given this preference score, however, it may not always be
possible to select a global ranking for all the knights. For instance,
Jamie Lannister may have ranked higher than Barristan Selmy in more
tournaments, and Barristan Selmy may have ranked higher than Loras
Tyrell in more tournaments but Loras Tyrell may still have ranked
higher than Jamie Lannister in more tournaments. For instance,
consider the following outcomes for the three knights in three
tournaments:

\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
& King's Landing Tourney  & Winterfell Tourney  & Casterly Rock
Tourney\\
\hline
Jamie (J) & 1$^{st}$ & 2$^{nd}$ & 3$^{rd}$\\
Barristan (B) & 2$^{nd}$ & 3$^{rd}$ & 1$^{st}$\\
Loras (L) & 3$^{rd}$ & 1$^{st}$ & 2$^{nd}$\\
\hline
\end{tabular}
\end{center}

\noindent In this case, $d(J,B) = d(B,L) = d(L,J) = 2$, and we have a
rock/paper/scissors situation in terms of preference: Jamie >
Barristan > Loras > Jamie. \\

\noindent To address such issues and provide a global ranking for the
knights given their rankings in individual events, the idea of the
Grand Maester and the Citadel is the following:\\

\noindent For each pair of knights $A$ and $B$ it is possible to
identify a ``preference sequence'' $s[A,B]$ of knights that might
imply that $A$ is more powerful than $B$. Such a sequence of size $m$
has the following properties:
\begin{itemize}
\item $s_1 = A$ and $s_m = B$
\item $\forall\ i \in [1,m-1]:$ $d[s_i,s_{i+1}] > d[s_{i+1},s_i]$
\end{itemize}
\noindent The ``strength of a preference sequence is defined as the
minimum preference score along the sequence: $$ t( s[A,B] ) =
\min_{\forall i \in [1,m-1]} d[s_i,s_{i+1}].$$ 

\noindent The ``strongest preference sequence'' $s^*[A,B]$ is the
preference sequence between the two knights $A$ and $B$ that has the
maximum strength:
$$s^*[A,B] = argmax_{\forall s[A,B]} t( s[A,B] ).$$ 

\noindent The strength of the strongest preference sequence between
two candidates is denoted as $t^*[A,B]$. If there is no preference
sequence between candidates A and B, then: $t^*[A,B] = 0$. It can be
shown that $t^*$ satisfies the transitive property (i.e., if $t^*[X,Y]
> t^*[Y,X]$ and $t^*[Y,Z] > t^*[Z,Y]$, then $t^*[X,Z] >
t^*[Z,X]$). Then it is possible to define a global preference function
between two knights, as follows:\vspace{-.2in}

\begin{equation*}
T(A,B)=\left\{
\begin{matrix}+1\ &\ if\ t^*[A,B]>t^*[B,A]\\
-1\ &\ if\ t^*[A,B]<t^*[B,A]\\
0\ &\ otherwise
\end{matrix}
\right\}
\end{equation*}

\noindent Given the Grand Maester's approach, there is at least one
knight $X$ that has better or equal global preference versus all other
knights $Y$, i.e., $\forall\ Y: T(X,Y) = 1$ or $T(X,Y)=0$. Then knight
$X$ is the ``global winner'' of all the tournaments and becomes the
Lord Commander of the Kingsguard.  In order to generate the global
ranking, the approach of the Grand Maester is to compute the strength
of the strongest preference sequence for every pair of knights and
rank the knights by finding first the Lord Commander, then removing
him from the list, finding the next winner, etc.\\

\noindent \textbf{A)} The Grand Maester argues that he has an
efficient algorithm for computing the reliability of the strongest
preference sequences for every pair of knights. What do you believe is
this efficient algorithm? In this process, describe the underlying
data structure needed to compute these values. What is the running
time of the efficient solution?\\

\noindent \textbf{B)} Assume that there are 5 knights that
participated in the tournaments: Barristan Selmy, Sandor Clegane,
Loras Tyrell, Brienne of Tarth and Jamie Lannister. There were 45
different tournaments that took place:\\

\noindent In 8 tournaments the ranking was:\\ Jamie Lannister > Brienne of Tarth
> Sandor Clegane > Loras Tyrell > Barristan Selmy\\

\noindent In 5 tournaments the ranking was:\\ Loras Tyrell > Barristan Selmy >
Jamie Lannister > Brienne of Tarth > Sandor Clegane\\

\noindent In 7 tournaments the ranking was:\\ Sandor Clegane > Barristan Selmy >
Brienne of Tarth > Jamie Lannister > Loras Tyrell\\

\noindent In 2 tournaments the ranking was:\\ Barristan Selmy > Jamie Lannister >
Loras Tyrell > Sandor Clegane > Brienne of Tarth\\

\noindent In 8 tournaments the ranking was:\\ Brienne of Tarth > Jamie Lannister >
Loras Tyrell > Sandor Clegane > Barristan Selmy\\

\noindent In 5 tournaments the ranking was:\\ Loras Tyrell > Sandor Clegane >
Brienne of Tarth > Barristan Selmy > Jamie Lannister\\

\noindent In 7 tournaments the ranking was:\\ Barristan Selmy > Loras Tyrell >
Brienne of Tarth > Jamie Lannister > Sandor Clegane\\

\noindent In 3 tournaments the ranking was:\\ Barristan Selmy > Loras Tyrell >
Jamie Lannister > Brienne of Tarth > Sandor Clegane\\

\noindent i) Compute all the preference values. Does an undisputed winner
arise in this case and why yes or no?\\

\noindent ii) Compute the strengths of the strongest preference
sequences for all pairs of knights with the efficient algorithm of the
Grand Maester.\\

\noindent iii) Provide the final ranking according to the strengths of
the strongest preference sequences.\\


\end{document}

